Discrete Quantum Theories and Computing

### 2018-11-25 dissertation.tex
dissertation.lyx "2.2 Quantum Probability" Updating...

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events \cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}.
Although they can be used to describe an individual quantum experiment,
to describe a family of quantum experiments, we would like to glue
the corresponding family of classical probability space together to
define a quantum probability space. The glued quantum probability
space is well-behaved since not only we can define the expectation
values, but the quantum probability measure define on the whole space
can be simply induced by the Born rule according to Gleason's theorem
\cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.

@@

Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps an event~$E$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcome in $E$ will actually happen. To maintain
the coherence, $\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, and $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
where $\overline{E}$ is the complement of $E$.  Moreover, we require
$\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

@@

One special feature of quantum experiments is that the probabilities
in different experiments are correlated. Consider another experiment
sending exactly the same particle as the previous one, but using a
different beam splitter corresponding to the observable $\mathbf{O}_{1}$
with spectral decomposition $\mathbf{O}_{1}=0\,\proj{\ps}+1\,\proj{\ms}+2\,\proj{2}$,
where 
\begin{align}
\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}\,, &  & \ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}\,.\label{eq:psms}
\end{align}
Although the sample space $\Omega_{1}=\left\{ \ket{\ps},\ket{\ms},\ket{2}\right\} $
and the probability measure $\mu_{1}:2^{\Omega_{1}}\rightarrow\left[0,1\right]$
defined by this experiment is different from the previous one, these
two experiments may produce the same post-experimental state~$\ket{2}$,
and the probability of the common event $\left\{ \ket{2}\right\} $
is believed to be the same, i.e., $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$.
In general, as long as sending the same particle, the probability
of the same event in different experiments should always be the same.
This fact is equivalent to the fact that commuting observables could
be measured simultaneously which is essential to define contextuality
and will be explained in Sec.~\ref{sec:Kochen-Specker}.

Since the probability induced by the different beam splitters are
correlated, it is more natural to define one quantum event space~$\events$
containing all possible classical event spaces using different beam
splitters. However, simply taking the union of all event spaces is
a bad idea because two events might appear different but representing
the same situation. For example, if we take the complement on the
both sides of $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$,
we will have
\begin{equation}
\mu_{0}\left(\left\{ \ket{0},\ket{1}\right\} \right)=1-\mu_{0}\left(\left\{ \ket{2}\right\} \right)=1-\mu_{1}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,,
\end{equation}
that is, the probabilities of events $\left\{ \ket{0},\ket{1}\right\} $
and $\left\{ \ket{\ps},\ket{\ms}\right\} $ are always the same, and
these events should be identified as the same quantum event to simplify
our discussion. This can be achieved by mapping a classical event~$E$
to the projector generated by $E$
\begin{equation}
\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\label{eq:pullback-function}
\end{equation}
with the convention $\varphi\left(\emptyset\right)=\mathbb{0}$ because
\begin{equation}
\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)=\proj{0}+\proj{1}=\proj{\ps}+\proj{\ms}=\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,.
\end{equation}
In general, if two classical events~$E$ and $E'$ mapped the same
projector, i.e., $\varphi\left(E\right)=\varphi\left(E'\right)$,
then the probability of $E$ is the same as the probability of $E'$.
Because of these good properties, the projector $\varphi\left(E\right)$
is called the quantum event corresponding to $E$, and the set of
all projectors on a given Hilbert space is called a quantum event
space~$\events$.

This function $\varphi$ also naturally sends the set structure to
the corresponding projector structure: $\varphi\left(\Omega\right)=\mathbb{1}$
and $\varphi\left(\overline{E}\right)=\mathbb{1}-\varphi\left(E\right)$.
Given two \emph{commuting} projectors $P_{0}$ and~$P_{1}$, there
exists a sample space $\Omega$ and a pair of event $E_{0}\subseteq\Omega$
and $E_{1}\subseteq\Omega$ such that $P_{0}=\varphi\left(E_{0}\right)$
and $P_{1}=\varphi\left(E_{1}\right)$. Conversely, given a pair of
event $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$, their corresponding
quantum events $\varphi\left(E_{0}\right)$ and $\varphi\left(E_{1}\right)$
are commuting and satisfying the following properties: 
\begin{align}
\varphi\left(E_{0}\cap E_{1}\right) & =\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,, & \varphi\left(E_{0}\cup E_{1}\right) & =\varphi\left(E_{0}\right)+\varphi\left(E_{1}\right)-\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,.\label{eq:pullback-properties}
\end{align}
Moreover, $E_{0}$ and $E_{1}$ are disjoint if and only if $\varphi\left(E_{0}\right)$
and $\varphi\left(E_{1}\right)$ are orthogonal, where two projectors
$P_{0}$ and $P_{1}$ are called \emph{orthogonal} if $P_{0}P_{1}=\mathbb{0}$.

Then, a quantum probability space can be defined by a quantum event
space~$\events$ together with a quantum probability measure $\mu:\events\rightarrow\left[0,1\right]$
subject to the corresponding constraints \cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010,Abramsky2012}:
\begin{align}
\mu\left(\mathbb{0}\right) & =0\,, & \mu\left(\mathbb{1}\right) & =1\,, & \mu\left(\mathbb{1}-P\right) & =1-\mu\left(P\right)\,,\label{eq:QuantumProbability}
\end{align}
and for each pair of orthogonal projectors $P_{0}$ and $P_{1}$:
\begin{equation}
\mu\left(P_{0}+P_{1}\right)=\mu\left(P_{0}\right)+\mu\left(P_{1}\right)\,.\label{eq:QuantumProbability-Addition}
\end{equation}
By defining in this way, given any orthonormal basis~$\Omega$, the
restricted function $\varphi:2^{\Omega}\rightarrow\events$ can pull
$\mu$ back to a classical probability measure $\varphi^{*}\mu:2^{\Omega}\rightarrow\left[0,1\right]$
by precomposition $\left(\varphi^{*}\mu\right)\left(E\right)=\mu\left(\varphi\left(E\right)\right)$.

@@

After introducing quantum probability measures, we might follow the
convention to introduce the Born rule which is the only way corresponding
a state to construct quantum probability measures in CQT. However,
since the variants of quantum probability measures in Sec.~\ref{sec:Toward-IVPM}
and Chapter~\ref{chap:QIVPM} could not be constructed by the Born
rule easily, it is instructive to search for a Born rule here, and
we will actually find the unique one.

Although quantum probability measures constructed by gluing together
the classical ones looks complex, they could all be induced by operators
according to Gleason's theorem.

@@

Although Gleason's theorem can correspond every quantum probability
measure to a unique mixed state~$\rho$, it doesn't explain whether
this state $\rho$ characterizing the state of particle sending to
a Stern-Gerlach type experiment. For example, in contrast with CQT,
if a quantum theory always ignore the input state and predict the
experimental outcomes are equally probable, even if this quantum theory
maps every input state to a quantum probability measure, it is still
not very sensible.

Let a pure unnormalized state $\ket{\Phi}\in\mathcal{H}$ characterize
the particle sending to a Stern-Gerlach type experiment, and $\muB_{\Phi}$
be the corresponding quantum probability measure of the resulting
events. To define a sensible correspondence, we hope that if the state~$\ket{\Phi}$
is one of the outcome of a quantum event~$P$, i.e., $P\ket{\Phi}=\ket{\Phi}$,
then the event~$P$ always happens, $\muB_{\Phi}\left(P\right)=1$
and vice versa. Moreover, since the physical phenomena exists and
should be the same no matter how we describe them, the probability
of an event should be invariant despite how we choose the basis. Because
changing to another basis is the same as applying an unitary map~$U$,
we should have$\muF_{U\ket{\Phi}}\left(UPU^{\dagger}\right)=\muF_{\Phi}\left(P\right)$.
It is easy to check the correspondence satisfying the conditions in
this paragraph is unique, 
\begin{equation}
\muB_{\Phi}(P)=\frac{\melem{\Phi}{P}{\Phi}}{\ip{\Phi}{\Phi}}\label{eq:Born}
\end{equation}
and called the Born rule \cite{Born1983,Mermin2007,Jaeger2007}. If
$\ket{\Phi}$ is normalized, Eq.~(\ref{eq:Born}) could be simplified
as $\muB_{\Phi}(P)=\melem{\Phi}{P}{\Phi}$. For mixed states $\rho=\sum_{j=1}^{N}q_{j}\proj{\Phi_{j}}$,
where $\ket{\Phi_{j}}\in\mathcal{H}$ are normalized, $q_{j}>0$,
and $\sum_{j=1}^{N}q_{j}=1$, the generalized Born rule induces a
probability measure $\muB_{\rho}$ defined as 
\begin{equation}
\muB_{\rho}\left(P\right)=\Tr\left(\rho P\right)=\sum_{j=1}^{N}q_{j}\muB_{\Phi_{j}}\left(P\right)\,.\label{eq:mixed-Born}
\end{equation}

As an example, consider a three-dimensional Hilbert space with orthonormal
basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $ and an observable
$\mathbf{O}_{0}$ defined in Eq.~(\ref{eq:observable}). Two fragments
of valid probability measures~$\mu_{1}$ and $\mu_{2}$ that can
be associated with this space are defined in Table~\ref{tab:quantum-probability-measure}.
By the Born rule, the first probability measure corresponds to the
quantum system being in the pure state $\ket{\ps}$ defined in Eq.~(\ref{eq:psms})
and the second corresponds to the quantum system being in the state
$\frac{\proj{0}+\proj{2}}{2}$. The expectation values of the observable
$\mathbf{O}$, $\expval{\mathbf{O}}_{\mu_{1,2}}$, are $1.5$ in the
first case and $2$ in the second. The quantum expectation value can
also used to decide whether a state is entangled or not for multipartite
systems as we describe in the following section.
\begin{table}
\noindent \begin{centering}
\caption{\label{tab:quantum-probability-measure}Two fragments of valid probability
measures~$\mu_{1}$ and $\mu_{2}$.}
\par\end{centering}
\noindent \centering{}%
\begin{tabular}{ccccccccccc}
\toprule 
\addlinespace
$\ket{\Psi}$ & $\ket{0}$ & $\ket{1}$ & $\ket{2}$ & $\frac{\ket{0}+\ket{1}}{\sqrt{2}}$ & $\frac{\ket{0}+\rmi\ket{1}}{\sqrt{2}}$ & $\frac{\ket{0}+\ket{2}}{\sqrt{2}}$ & $\frac{\ket{0}+\rmi\ket{2}}{\sqrt{2}}$ & $\frac{\ket{1}+\ket{2}}{\sqrt{2}}$ & $\frac{\ket{1}+\rmi\ket{2}}{\sqrt{2}}$ & \ldots\tabularnewline\addlinespace
\midrule
\addlinespace
\addlinespace
$\mu_{1}\left(\proj{\Psi}\right)$ & $\frac{1}{2}$ & $\frac{1}{2}$ & $0$ & $1$  & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & \ldots\tabularnewline\addlinespace
\addlinespace
\addlinespace
$\mu_{2}\left(\proj{\Psi}\right)$ & $\frac{1}{2}$ & $0$ & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{2}$ & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & \ldots\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\end{table}

>>>>

A \emph{probability space} is a mathematical abstraction specifying
the necessary conditions for reasoning coherently about collections
of uncertain events \cite{Kolmogorov1950,544199,Griffiths2003,Grabisch2016}.
Although they can be used to describe an individual quantum experiment,
to describe a family of quantum experiments, we would like to glue
their probability spaces together to define a quantum probability
space. The glued quantum probability space is well-behaved since not
only we can define the expectation values, but the quantum probability
measure define on the whole space can be simply induced by the Born
rule according to Gleason's theorem \cite{gleason1957,Redhead1987-REDINA,peres1995quantum,RichmanBridges1999,Hamhalter2013}.

@@

Given a finite sample space~$\Omega$ representing all possible outcomes
of a process, and its power set $2^{\Omega}$ as the classical event
space, a classical probability measure~$\mu$ maps an event~$E$
to a number $\mu\left(E\right)\in\left[0,1\right]$ specifying how
likely one of the outcomes in $E$ will actually happen. To maintain
the coherence, $\mu$ is subject to the following constraints: $\mu\left(\emptyset\right)=0$,
$\mu\left(\Omega\right)=1$, and $\mu\left(\overline{E}\right)=1-\mu\left(E\right)$,
where $\overline{E}$ is the complement of $E$.  Moreover, we require
$\mu\left(E_{0}\cup E_{1}\right)=\mu\left(E_{0}\right)+\mu\left(E_{1}\right)$
for each pair of disjoint events $E_{0}\subseteq\Omega$ and $E_{1}\subseteq\Omega$.

@@

One special feature of quantum experiments is that the probabilities
in different experiments are correlated. Consider another experiment
sending exactly the same particle as the previous one, but using a
different beam splitter corresponding to the observable $\mathbf{O}_{1}$
with spectral decomposition $\mathbf{O}_{1}=0\,\proj{\ps}+1\,\proj{\ms}+2\,\proj{2}$,
where $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$ and $\ket{\ms}=\frac{\ket{0}-\ket{1}}{\sqrt{2}}$.
Although the sample space $\Omega_{1}=\left\{ \ket{\ps},\ket{\ms},\ket{2}\right\} $
and the probability measure $\mu_{1}:2^{\Omega_{1}}\rightarrow\left[0,1\right]$
defined by this experiment is different from the previous one, these
two experiments may produce the same post-experimental state~$\ket{2}$,
and the probability of the common event $\left\{ \ket{2}\right\} $
is believed to be the same, i.e., $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$.
In general, as long as sending the same particle, the probability
of the same event in different experiments should always be the same.
This fact is equivalent to the fact that commuting observables could
be measured simultaneously which is essential to define contextuality
and will be explained in Sec.~\ref{sec:Kochen-Specker}.

Since the probability induced by the different beam splitters are
correlated, it is more natural to define one quantum event space~$\events$
containing all possible classical event spaces using different beam
splitters. However, simply taking the union of all event spaces is
a bad idea because two events might appear different but representing
the same situation. For example, if we take the complement on the
both sides of $\mu_{0}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{2}\right\} \right)$,
we will have
\begin{equation}
\mu_{0}\left(\left\{ \ket{0},\ket{1}\right\} \right)=1-\mu_{0}\left(\left\{ \ket{2}\right\} \right)=1-\mu_{1}\left(\left\{ \ket{2}\right\} \right)=\mu_{1}\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,,
\end{equation}
that is, the probabilities of events $\left\{ \ket{0},\ket{1}\right\} $
and $\left\{ \ket{\ps},\ket{\ms}\right\} $ are always the same, and
these events should be identified as the same quantum event to simplify
our discussion. This can be achieved by mapping a classical event~$E$
to the projector generated by $E$
\begin{equation}
\varphi\left(E\right)=\sum_{\ket{j}\in E}\proj{j}\label{eq:pullback-function}
\end{equation}
with the convention $\varphi\left(\emptyset\right)=\mathbb{0}$ because
\begin{equation}
\varphi\left(\left\{ \ket{0},\ket{1}\right\} \right)=\proj{0}+\proj{1}=\proj{\ps}+\proj{\ms}=\varphi\left(\left\{ \ket{\ps},\ket{\ms}\right\} \right)\,.
\end{equation}
In general, if two classical events~$E$ and $E'$ are mapped to
the same projector, i.e., $\varphi\left(E\right)=\varphi\left(E'\right)$,
then the probability of $E$ is the same as the probability of $E'$.
Therefore, for any classical event~$E$, its corresponding quantum
event is defined to be the projector $\varphi\left(E\right)$, and
the set of all projectors on a given Hilbert space is called a quantum
event space~$\events$.

This function $\varphi$ not only respects the probability of events
but also naturally sends the set structure to the corresponding projector
structure: $\varphi\left(\Omega\right)=\mathbb{1}$ and $\varphi\left(\overline{E}\right)=\mathbb{1}-\varphi\left(E\right)$.
Given two \emph{commuting} projectors $P_{0}$ and~$P_{1}$, there
exists a sample space $\Omega$ and a pair of events $E_{0}\subseteq\Omega$
and $E_{1}\subseteq\Omega$ such that $P_{0}=\varphi\left(E_{0}\right)$
and $P_{1}=\varphi\left(E_{1}\right)$. Conversely, given a pair of
events $E_{0}$ and $E_{1}$ in the same sample space~$\Omega$,
their corresponding quantum events $\varphi\left(E_{0}\right)$ and
$\varphi\left(E_{1}\right)$ are commuting and satisfying the following
properties: 
\begin{align}
\varphi\left(E_{0}\cap E_{1}\right) & =\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,, & \varphi\left(E_{0}\cup E_{1}\right) & =\varphi\left(E_{0}\right)+\varphi\left(E_{1}\right)-\varphi\left(E_{0}\right)\varphi\left(E_{1}\right)\,.\label{eq:pullback-properties}
\end{align}
Moreover, $E_{0}$ and $E_{1}$ are disjoint if and only if $\varphi\left(E_{0}\right)$
and $\varphi\left(E_{1}\right)$ are orthogonal, where two projectors
$P_{0}$ and $P_{1}$ are called \emph{orthogonal} if $P_{0}P_{1}=\mathbb{0}$.

Then, a quantum probability space can be defined by a quantum event
space~$\events$ together with a quantum probability measure $\mu:\events\rightarrow\left[0,1\right]$
subject to the corresponding constraints \cite{10.2307/2308516,gleason1957,Redhead1987-REDINA,Maassen2010,Abramsky2012}:
\begin{align}
\mu\left(\mathbb{0}\right) & =0\,, & \mu\left(\mathbb{1}\right) & =1\,, & \mu\left(\mathbb{1}-P\right) & =1-\mu\left(P\right)\,,\label{eq:QuantumProbability}
\end{align}
and for each pair of orthogonal projectors $P_{0}$ and $P_{1}$:
\begin{equation}
\mu\left(P_{0}+P_{1}\right)=\mu\left(P_{0}\right)+\mu\left(P_{1}\right)\,.\label{eq:QuantumProbability-Addition}
\end{equation}
Because $\varphi$ respects the probability of events, if we restrict
$\varphi$ on an event space $2^{\Omega}$, the function $\varphi^{*}\mu:2^{\Omega}\rightarrow\left[0,1\right]$
defined by precomposition $\left(\varphi^{*}\mu\right)\left(E\right)=\mu\left(\varphi\left(E\right)\right)$
is a classical probability measure and called the pullback of $\mu$
by $\varphi:2^{\Omega}\rightarrow\events$.

@@

After introducing quantum probability measures, we might follow the
convention to introduce the Born rule which is the only way corresponding
a state to a quantum probability measure in CQT. However, since the
variants of quantum probability measures in Sec.~\ref{sec:Toward-IVPM}
and Chapter~\ref{chap:QIVPM} could not be constructed by the Born
rule easily, searching for a Born rule step-by-step here might provide
a idea what we should do in Sec.~\ref{sec:Toward-IVPM} and Chapter~\ref{chap:QIVPM}.
While the situations will become more delicate in the later sections,
we will fortunately find the unique Born rule in CQT here.

Although a quantum probability measure constructed by gluing together
the classical ones looks complex, it could be induced by an operator
according to Gleason's theorem.

@@

Although Gleason's theorem can correspond every quantum probability
measure~$\mu$ to a unique mixed state~$\rho$, if we follow the
Stern-Gerlach type experiment interpretation in previous section,
we can find Gleason's theorem doesn't specify whether this state $\rho$
characterizing the state of particle sending to the quantum experiment.
For example, in contrast with CQT, a quantum theory could basically
ignore the input state and predict the experimental outcomes are equally
probable~\cite{doi:10.1142/S0217984913500644,1751-8121-47-40-405304}.
Even if a version of Bell's inequality could be proved in this quantum
theory, its prediction might be different from the prediction of CQT
enough that it is easily distinguish them experimentally.

Let a pure unnormalized state $\ket{\Phi}\in\mathcal{H}$ characterize
the particle sending to a Stern-Gerlach type experiment, and $\muB_{\Phi}$
be the corresponding quantum probability measure of the resulting
events. To define a sensible correspondence, we hope that if the state~$\ket{\Phi}$
is one of the outcomes of a quantum event~$P$, i.e., $P\ket{\Phi}=\ket{\Phi}$,
then the event~$P$ always happens, $\muB_{\Phi}\left(P\right)=1$
and vice versa. Moreover, since the physical phenomena exists and
should be the same no matter how we describe them, the probability
of an event should be invariant despite how we choose the basis. Because
changing to another basis is the same as applying an unitary map~$U$,
we should have $\muF_{U\ket{\Phi}}\left(UPU^{\dagger}\right)=\muF_{\Phi}\left(P\right)$.
It is easy to check that the correspondence satisfying the conditions
in this paragraph is unique, 
\begin{equation}
\muB_{\Phi}(P)=\frac{\melem{\Phi}{P}{\Phi}}{\ip{\Phi}{\Phi}}\label{eq:Born}
\end{equation}
and called the Born rule \cite{Born1983,Mermin2007,Jaeger2007}. If
$\ket{\Phi}$ is normalized, Eq.~(\ref{eq:Born}) could be simplified
as $\muB_{\Phi}(P)=\melem{\Phi}{P}{\Phi}$. For mixed states $\rho=\sum_{j=1}^{N}q_{j}\proj{\Phi_{j}}$,
where $\ket{\Phi_{j}}\in\mathcal{H}$ are normalized, $q_{j}>0$,
and $\sum_{j=1}^{N}q_{j}=1$, the generalized Born rule induces a
probability measure $\muB_{\rho}$ defined as 
\begin{equation}
\muB_{\rho}\left(P\right)=\Tr\left(\rho P\right)=\sum_{j=1}^{N}q_{j}\muB_{\Phi_{j}}\left(P\right)\,.\label{eq:mixed-Born}
\end{equation}

As an example, consider a three-dimensional Hilbert space with orthonormal
basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $ and an observable
$\mathbf{O}_{0}$ defined in Eq.~(\ref{eq:observable}). Two fragments
of valid probability measures~$\mu_{1}$ and $\mu_{2}$ that can
be associated with this space are defined in Table~\ref{tab:quantum-probability-measure}.
By the Born rule, the first probability measure corresponds to the
quantum system being in the pure state $\ket{\ps}=\frac{\ket{0}+\ket{1}}{\sqrt{2}}$
and the second corresponds to the quantum system being in the state
$\frac{\proj{0}+\proj{2}}{2}$. The expectation values of the observable
$\mathbf{O}$, $\expval{\mathbf{O}}_{\mu_{1,2}}$, are $1.5$ in the
first case and $2$ in the second. The quantum expectation value can
also used to decide whether a state is entangled or not for multipartite
systems as we describe in the following section.
\begin{table}
\noindent \begin{centering}
\caption{\label{tab:quantum-probability-measure}Two fragments of valid probability
measures~$\mu_{1}$ and $\mu_{2}$.}
\par\end{centering}
\noindent \centering{}%
\begin{tabular}{ccccccccccc}
\toprule 
\addlinespace
$\ket{\Psi}$ & $\ket{0}$ & $\ket{1}$ & $\ket{2}$ & $\frac{\ket{0}+\ket{1}}{\sqrt{2}}$ & $\frac{\ket{0}+\rmi\ket{1}}{\sqrt{2}}$ & $\frac{\ket{0}+\ket{2}}{\sqrt{2}}$ & $\frac{\ket{0}+\rmi\ket{2}}{\sqrt{2}}$ & $\frac{\ket{1}+\ket{2}}{\sqrt{2}}$ & $\frac{\ket{1}+\rmi\ket{2}}{\sqrt{2}}$ & \ldots\tabularnewline\addlinespace
\midrule
\addlinespace
\addlinespace
$\mu_{1}\left(\proj{\Psi}\right)$ & $\frac{1}{2}$ & $\frac{1}{2}$ & $0$ & $1$  & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & \ldots\tabularnewline\addlinespace
\addlinespace
\addlinespace
$\mu_{2}\left(\proj{\Psi}\right)$ & $\frac{1}{2}$ & $0$ & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & $\frac{1}{2}$ & $\frac{1}{2}$ & $\frac{1}{4}$ & $\frac{1}{4}$ & \ldots\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabular}
\end{table}
